---
title: "Camera traps Wildlife"
author:
  - Osa Conservation Wildlife Team and Earthwatch
site: bookdown::bookdown_site
output:
  bookdown::gitbook:
    css: styles.css
date: '`r paste("Updated on", Sys.Date())`'
documentclass: book
biblio-style: apalike
link-citations: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Wildlife is one of the programs that Osa Consevation harbours, you can hace a look of the organizacion here: https://osaconservation.org/


The main goal of Wildlife program is to understand, protect and restore wildlife within the Amistosa climate lifeboat for future generations through biodiversity monitoring at the landscape scale, application of conservation technologies, engaging local stakeholders, and adopting evidence-based restoration techniques, ultimately advancing climate change adaptation and resilience of the South Pacific region of Costa Rica.

One of the main tools used here are camera traps. Those are automa

Here we provide the main results of the current projects carried out by our team.


## Study area and survey locations

```{r c01, echo=F, message=F, include=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)

library(dplyr)
library(tidyr)
library(leaflet)
library(units)
library(sf)
library(viridis)
library(kableExtra)
library(lubridate)
library(plotly)
library(googledrive)
library(purrr)
library(readxl)
library(RColorBrewer)
library(scales)
library(ggrepel)
library(stringr)


#drive_auth()
googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
data_path <- "data"
dir.create(data_path) 
#test2
# For when the trapping effort file is sorted
# googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))

summary(cars)
```

# Long Term Monitoring

You can also embed plots, for example:

```{r, c02, include=F, message=F, warning=F, results="hide"}

#Import the shapefiles for the maps
dir.create("shapefiles")
shp_folder <- drive_ls(path = as_id("1tIPtT8gXcLRPW26IqH1wOjm9_cjhJG9L"), recursive=T)

drive_download(file = as_id(shp_folder$id[2]), path = paste0("shapefiles/", shp_folder$name[2]), overwrite = TRUE)

LTM_Map <- st_read("shapefiles/LongTermMonitoring2026.kml")
LTM_Map2<- st_zm(LTM_Map, drop = TRUE, what = "ZM")
```


here goes the map

```{r}

leaflet(LTM_Map2) %>% 
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addProviderTiles(providers$OpenStreetMap, group = "Street Map") %>%
  addCircleMarkers(
    radius = 3, 
    fillOpacity = 0.8
  ) %>% 
  addLayersControl(
    baseGroups = c("Street Map","Satellite"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addScaleBar(position = "topright")

```

The following represents the data exploration for the data collected during the collaboration between Osa Conservation for the project entitled:

**Main Title.**

Also known as:

**Long term monitoring**

sdfjkhsdfl

# Exploration to Fila Costeña


# Arboreal Bridges



# Farms for Nature



# Piro Restoration plots
