---
title: "Camera traps Wildlife"
author:
  - Osa Conservation Wildlife Team and Earthwatch
site: bookdown::bookdown_site
output:
  bookdown::gitbook:
    css: styles.css
date: '`r paste("Updated on", Sys.Date())`'
documentclass: book
biblio-style: apalike
link-citations: yes
editor_options:
  chunk_output_type: console
---

# Introduction

Wildlife is one of the programs that Osa Consevation harbours, you can hace a look of the organizacion here: https://osaconservation.org/


The main goal of Wildlife program is to understand, protect and restore wildlife within the Amistosa climate lifeboat for future generations through biodiversity monitoring at the landscape scale, application of conservation technologies, engaging local stakeholders, and adopting evidence-based restoration techniques, ultimately advancing climate change adaptation and resilience of the South Pacific region of Costa Rica.

One of the main tools used here are camera traps. Those are automa

Here we provide the main results of the current projects carried out by our team.


## Study area and survey locations

```{r c01, echo=F, message=F, include=F}
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)

library(dplyr)
library(tidyr)
library(leaflet)
library(leaflet.extras)
library(units)
library(sf)
library(viridis)
library(kableExtra)
library(lubridate)
library(plotly)
library(googledrive)
library(purrr)
library(readxl)
library(RColorBrewer)
library(scales)
library(ggrepel)
library(stringr)


#drive_auth()
#googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))
#data_path <- "data"
#dir.create(data_path) 

#test2
# For when the trapping effort file is sorted
# googledrive::drive_auth(path = Sys.getenv("GOOGLE_AUTHENTICATION_CREDENTIALS"))

```

# Piro Restoration Plots


# Exploration to Fila Costeña


# Campus volunteers cameras


# Arboreal Bridges


# Farms for Nature


# Long Term Monitoring

*Introduction:* Loren ipsum ...


```{r, c02, include=F, message=F, warning=F, results="hide"}

#Import the shapefiles for the maps

#dir.create("shapefiles")
#shp_folder <- drive_ls(path = as_id("1tIPtT8gXcLRPW26IqH1wOjm9_cjhJG9L"), recursive=T)

#drive_download(file = as_id(shp_folder$id[2]), path = paste0("shapefiles/", shp_folder$name[2]), overwrite = TRUE)

LTM_Map <- st_read("shapefiles/LongTermMonitoring2026.kml")
LTM_Map2<- st_zm(LTM_Map, drop = TRUE, what = "ZM")

```

Here goes the map

```{r}

leaflet(LTM_Map2) %>% 
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addProviderTiles(providers$OpenStreetMap, group = "Street Map") %>%
  addCircleMarkers(
    radius = 3, 
    fillOpacity = 0.8
  ) %>% 
  addLayersControl(
    baseGroups = c("Street Map","Satellite"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addScaleBar(position = "topright")

```

The following represents the data exploration for the data collected for the project:

``` {r}
# Download the Wildlife Insights data and format it
#WildlifeInsights_folder <- drive_ls(path = as_id("1L-NrbOSaqtxzHrukcuxZAXSg4Ate35iA"), recursive=T)

#download images
#drive_download(file = as_id(WildlifeInsights_folder$id[2]), path = WildlifeInsights_folder$name[2], overwrite = TRUE)

#Download deployments
#drive_download(file = as_id(WildlifeInsights_folder$id[5]), path =  WildlifeInsights_folder$name[5], overwrite = TRUE)

img0 <- read.csv("data/images_2003884.csv", header=T)
dep0 <- read.csv("data/deployments.csv", header=T)

#Subset Long Term Monitoring
dep<-dep0 %>% 
  filter(subproject_name=="Long Term Monitoring")

img <- img0 %>% 
  filter(deployment_id %in% dep$deployment_id)

#
#Subset Fila Costeña
#dep<-dep0 %>% 
#  filter(str_starts(subproject_name, "Fila"))

#img <- img0 %>% 
#  filter(deployment_id %in% dep$deployment_id)


#Subset 
#dep<-dep0 %>% 
#  filter(str_starts(subproject_name, "R2R"))

#img <- img0 %>% 
#  filter(deployment_id %in% dep$deployment_id)



# Prep the dates 

# start dates
dep$start_date <- ymd_hms(paste0(substr(dep$start_date,1,10), " 00:00:01"))

# end dates
dep$end_date   <- ymd_hms(paste0(substr(dep$end_date,1,10), " 23:59:00"))

# camera days
dep$days <- round(interval(dep$start_date, dep$end_date)/ddays(1),0)
#dep$feature_type <- substr(dep$placename,4,4)

# Image dates
# Image timestamp
img$timestamp <- ymd_hms(img$timestamp)
#
dep <- dep[duplicated(dep[,c("deployment_id", "placename")])==F,]
img <- left_join(img, dep[,c("deployment_id", "placename")])

```


To date we have deployed `r nrow(dep)` cameras, at `r length(unique(dep$placename))` locations. Resulting in the capture of `r nrow(img[img$is_blank==0,])` wildlife images of `r length(unique(img$common_name))` species.


```{r, cam_dat}

# Create the leaflet map
LocationsMap <- leaflet(data = dep) %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addProviderTiles(providers$OpenStreetMap, group = "Street Map")  %>%
  addCircleMarkers(
    radius = 3, 
    fillOpacity = 0.8
  ) %>%
  addLayersControl(
    position = "topleft",
    baseGroups = c("Satellite", "Street Map"),
    options = layersControlOptions(collapsed = FALSE)
  )

LocationsMap

```


Fist take a look of the chronology of the deployments:

```{r}
# Call the plot
ChronoPlot <- plot_ly()

# We want a separate row for each 'placename' - so lets turn it into a factor
dep$placename <- as.factor(dep$placename)

# loop through each place name
for(i in seq_along(levels(dep$placename)))
  {
      #Subset the data to just that placename
      tmp <- dep[dep$placename==levels(dep$placename)[i],]
      # Order by date
      tmp <- tmp[order(tmp$start_date),]
      # Loop through each deployment at that placename
      for(j in 1:nrow(tmp))
      {
        # Add a line to 'p'
        ChronoPlot <- add_trace(ChronoPlot, 
                       #Use the start and end date as x coordinates
                       x = c(tmp$start_date[j], tmp$end_date[j]), 
                       #Use the counter for the y coordinates
                       y = c(i,i), 
                       # State the type of chart
                       type="scatter",
                       # make a line that also has points
                       mode = "lines+markers", 
                       # Add the deployment ID as hover text
                       hovertext=tmp$deployment_id[j], 
                       # Color it all black
                       color=I("black"), 
                       # Suppress the legend
                       showlegend = FALSE)
      }
      
  }
# Add a categorical y axis
 ChronoPlot <- ChronoPlot %>%   
   layout(yaxis = list(
      ticktext = as.list(levels(dep$placename)), 
      tickvals = as.list(1:length(levels(dep$placename))),
      tickmode = "array"))

ChronoPlot

#Filter down the image data to remove anything not identified to species:

#Camera activity plot
# Order by placename to make the graphs easier to follow
dep <- dep[order(dep$placename),]

# Get unique placenames
placenames_unique <- unique(dep$placename)

#Get a list of unique locations
locs <- dep[,c("placename", "longitude", "latitude")]
locs <- locs[duplicated(locs)==F,]

plot_height <- max(400, length(placenames_unique) * 20)


#Filter down the image data to remove anything not identified to species:
img_sub <- img %>% filter(is_blank==0,                # Remove the blanks
                          is.na(img$species)==FALSE, # Remove classifications which don't have species 
                          species!="sapiens",
                          species!="",
                          common_name!="No CV Result") 

img_subMammal<-img_sub %>% filter(class=="Mammalia")

img_subBird<-img_sub %>% filter(class=="Aves")

```


Which results in `r nrow(img_sub)` identifiable wildlife images of `r length(unique(img_sub$common_name))` species. Including `r length(unique(img_subMammal$common_name))` Mammals and `r length(unique(img_subBird$common_name))` Birds.

Now, we goint to check the results bases od independent records:


```{r}
# Prep the analysis dataframes

# Remove any deployments without end dates
depEnds <- dep[is.na(dep$end_date)==F,] #formerly tmp

# Create an empty list to store our days
daily_lookup <- list()

# Loop through the deployment dataframe and create a row for every day the camera is active
for(i in 1:nrow(depEnds))
{
  if(ymd(substr(depEnds$start_date[i],1,10))!=ymd(substr(depEnds$end_date[i],1,10)))
  {
    daily_lookup[[i]] <- data.frame(
      "date"=seq(ymd(substr(depEnds$start_date[i],1,10)),
                 ymd(substr(depEnds$end_date[i],1,10)), by="days"),
      "placename"=depEnds$placename[i])
  }
}

# Merge the lists into a dataframe
row_lookup <- bind_rows(daily_lookup)

# Remove duplicates - when start and end days are the same for successive deployments
row_lookup <- row_lookup[duplicated(row_lookup)==F,]

###################################
# Create your independent detections
independent <- 30

img_sub$animal_count <- img_sub$number_of_objects
img_sub$sp <- paste(img_sub$genus, img_sub$species, sep=".")

# Create your independent data

img_tmp <- img_sub %>%
              arrange(deployment_id) %>%        # Order by deployment_id
              group_by(deployment_id, sp) %>%   # Group species together
              mutate(duration = int_length(timestamp %--% lag(timestamp))) # Calculate the gap bet

# Give a random value to all cells
img_tmp$event_id <- 9999

# Create a counter
counter <- 1

# Make a unique code that has one more zero than rows in your dataframe  
num_code <- as.numeric(paste0(nrow(img_sub),0))

# Loop through img_tmp - if gap is greater than the threshold -> give it a new event ID
for (i in 2:nrow(img_tmp)) {
  img_tmp$event_id[i-1]  <- paste0("E", str_pad(counter, nchar(num_code), pad = "0"))
  
  if(is.na(img_tmp$duration[i]) | abs(img_tmp$duration[i]) > (independent * 60))
    {
      counter <- counter + 1
    }
}

# Update the information for the last row - the loop above always updates the previous row... leaving the last row unchanged
   
 # group ID  for the last row
 if(img_tmp$duration[nrow(img_tmp)] < (independent * 60)|
    is.na(img_tmp$duration[nrow(img_tmp)])){
   img_tmp$event_id[nrow(img_tmp)] <- img_tmp$event_id[nrow(img_tmp)-1]
 } else{
   counter <- counter + 1
   img_tmp$event_id[nrow(img_tmp)] <- paste0("E", str_pad(counter, nchar(num_code), pad = "0"))
 }

# remove the duration column
img_tmp$duration <- NULL
 
 # find out the last and the first of the time in the group
  top <- img_tmp %>% group_by(event_id) %>% top_n(1,timestamp) %>% dplyr::select(event_id, timestamp)
  bot <- img_tmp %>% group_by(event_id) %>% top_n(-1,timestamp) %>% dplyr::select(event_id, timestamp)
  names(bot)[2] <- c("timestamp_end")
  
  img_num <- img_tmp %>% 
    group_by(event_id) %>% 
    summarise(event_observations=n()) # number of images in the event
  
  event_grp <- img_tmp %>% 
    group_by(event_id) %>% 
    summarise(event_groupsize=max(animal_count))

  # calculate the duration and add the other elements
  diff <-  top %>% 
    left_join(bot, by="event_id") %>%
    mutate(event_duration=abs(int_length(timestamp %--% timestamp_end))) %>%
    left_join(event_grp, by="event_id")%>%
    left_join(img_num, by="event_id")

  # Remove columns you don't need
  diff$timestamp   <-NULL
  diff$timestamp_end <-NULL
  diff <- diff[duplicated(diff)==F,]
    # Merge the img_tmp with the event data
  img_tmp <- left_join(img_tmp,diff,by="event_id")
  
  
# Remove duplicates
ind_dat <- img_tmp %>% distinct(event_id, .keep_all = TRUE)

# Make a  unique code for ever day and deployment where cameras were functioning

#UniqueDayCode <- paste(row_lookup$date, row_lookup$placename) #previously tmp

#Subset ind_dat to data that matches the unique codes
#ind_dat2 <- ind_dat[paste(substr(ind_dat$timestamp,1,10), ind_dat$placename) %in% UniqueDayCode, ] #formerly ind_dat

# Convert your species names to factors
ind_dat$sp <- as.factor(ind_dat$sp)

#Subset the columns
dep_sites <- dep[, c("placename", "longitude", "latitude")] # formerly tmp

# Remove duplicated rows
dep_sites<- dep_sites[duplicated(dep_sites)==F,]

# Total counts
  # Station / Month / deport / Species      
  # Calculate the number of days at each site  
  total_obs <- row_lookup %>% 
      group_by(placename) %>%
      summarise(days = n())
  
  # Convert to a data frame
  total_obs <- as.data.frame(total_obs)
  
  # Add columns for each species  
  total_obs[, levels(ind_dat$sp)] <- NA
  
  # Duplicate for counts
  total_count <- total_obs
  # Test counter
  i <-1
  # For each station, count the number of individuals/observations
  for(i in 1:nrow(total_obs))
    {
      dep_sites <- ind_dat[ind_dat$placename==total_obs$placename[i],]
      
      tmp_stats <- dep_sites %>%  group_by(sp, .drop=F) %>% summarise(obs=n(), count=sum(animal_count))
      
      total_obs[i,as.character(tmp_stats$sp)] <- tmp_stats$obs
      total_count[i,as.character(tmp_stats$sp)] <- tmp_stats$count
    }

# add simple treatment to the total_count dataframe
total_count$plot_id <- substr(total_count$placename,4,6)  
#total_count <- left_join(total_count, dep$plot_id[, c("plot_id", #"simple_treatment")])  
  
```

This resulted in `r nrow(ind_dat)` independent detections of wildlife. The total number of detections and proportion of cameras detected at are as follows:

```{r}
# Make a species list
# First define vector of the headings you want to see (we will use this trick a lot later on)
taxonomy_headings <- c("class", "order", "family", "genus", "species", "sp", "common_name")

# Subset the image data to just those columns
ind_tax<- ind_dat[,colnames(ind_dat)%in% taxonomy_headings] #previously tmp
# Remove duplicates
ind_tax <- ind_tax[duplicated(ind_tax)==F,]

# Create an ordered species list
sp_summary0  <- ind_tax[order(ind_tax$class, ind_tax$order, ind_tax$family, ind_tax$genus, ind_tax$sp, ind_tax$species),]

# Convert to long
long_obs <- total_obs %>% 
  pivot_longer(cols=sp_summary0$sp,  # The columns we want to create into rows - species
               names_to="sp",       # What we what the number column to be called
               values_to = "count") # Takes the values in the species columns and calls them `count`

# We can them summaries those using dplyr
sppSummary <- long_obs %>%                   # Take the long observation data frame `long_obs` 
          group_by(sp) %>%            # Group by species
          summarise(count=sum(count)) # Sum all the independent observations

# Add it to the sp_summary dataframe
sp_summary <- left_join(sp_summary0, sppSummary)

## Occupancy
# We use the mutate function to mutate the column
total_binary <-  total_obs %>%    # The total obs dataframe              
                    mutate(across(sp_summary$sp, ~+as.logical(.x)))  # across all of the species columns, make it binary

# Flip the dataframe to longer - as before
long_bin <- total_binary %>% 
  pivot_longer(cols=sp_summary$sp, names_to="sp", values_to = "count") # Takes the species names columns, and makes them unique rows with "sp" as the key 

# We can now sum the presence/absences and divide by the number of survey locations
OccuSummary <- long_bin %>% 
  group_by(sp) %>% 
  summarise(occupancy=sum(count)/nrow(locs)) # divided the sum by the number of sites

# add the results to the sp_summary

sp_summary <- sp_summary[order(sp_summary$count),] #final


```


How is the distribution of Threatened species in our records?

```{r}

#drive_download(file = as_id("1MtXa2adHajT0hOuPNKd0JE2L-Qrp_U3T"), path = paste0("data/", "species_lists.xlsx"), overwrite = TRUE)

sp_mam <- read_excel("data/species_lists.xlsx", sheet="mammals_list") 
#sp_amp <- read_excel("data/species_lists.xlsx", sheet="amphibians_list") 
sp_bir <- read_excel("data/species_lists.xlsx", sheet="bird_list") 

# Merge the list

sp_all <- bind_rows(sp_mam[,c("scientific_name", "Endemism", "IUCN_status", "IUCN_population_trend", "class_name")],
#sp_amp[,c("scientific_name", "Endemism", "IUCN_status", "IUCN_population_trend", "class_name")],
sp_bir[,c("scientific_name", "Endemism", "IUCN_status", "IUCN_population_trend", "class_name")])

# Remove duplicates 
sp_all <- sp_all[duplicated(sp_all)==F,]


# Merge the species recorded on surveys
all_caps <- gsub("\\.", " ", sp_summary$sp)
iucn <- data.frame("scientific_name"=unlist(all_caps))

# Remove duplicated
iucn$tmp <- 1
iucn <- iucn[duplicated(iucn$scientific_name)==F,]
iucn$tmp <- NULL
iucn_all <- left_join(iucn, sp_all)

iucn_all$IUCN_status[iucn_all$IUCN_status=="Not assessed"] <- "Not Assessed"

#Order
iucn_all <- iucn_all[order(iucn_all$class_name, iucn_all$IUCN_status, iucn_all$Endemism, iucn_all$scientific_name),]


iucn_all<-iucn_all   %>%
  filter(!is.na(IUCN_status))

# Define IUCN colors including Not Assessed
iucn_colors <- c(
  "Least Concern" = "#90EE90",
  "Near Threatened" = "#FFA500",
  "Vulnerable" = "#FFFF00",
  "Endangered" = "#FF6B6B",
  "Critically Endangered" = "#8B0000",
  "Data Deficient" = "#999999",
  "Not Assessed" = "#EEEEEE",
  "Unknown" = "#FFFFFF"
)
############### WAFFLE PLOT

# Prepare data with grid positions
waffle_hover <- iucn_all %>%
  mutate(IUCN_status = ifelse(is.na(IUCN_status), "Not Assessed", IUCN_status)) %>%
  group_by(class_name) %>%
  mutate(
    # Create grid within each class
    row_num = (row_number() - 1) %/% 15 + 1,
    col_num = (row_number() - 1) %% 15 + 1,
    # Create hover text
    hover_text = paste(
      "Species:", scientific_name,
      "<br>Endemism:", Endemism,
      "<br>Population Trend:", IUCN_population_trend,
      "<br>IUCN Status:", IUCN_status
    )
  ) %>%
  ungroup()

# Create static ggplot
p <- ggplot(waffle_hover, aes(x = col_num, y = row_num, fill = IUCN_status, text = hover_text)) +
  geom_tile(color = "white", linewidth = 0.3) +
  facet_wrap(~class_name, scales = "fixed", ncol = 1) +
  scale_fill_manual(values = iucn_colors, name = "IUCN Status") +
  coord_equal() +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    legend.position = "bottom"
  ) +
  labs(
    title = "IUCN Status Distribution by Class",
    subtitle = "Each square represents one species. Hover for details."
  )

# Convert to interactive plotly
ggplotly(p, tooltip = "text") %>%
  layout(
    hoverlabel = list(bgcolor = "white", font = list(size = 11)),
    legend = list(orientation = "h", y = -0.1)
  )


```


Check the number of records and the occupancy

```{r}

###########################
# Comparison plot
# Lets put the dataframes in a sensible order

yform <- list(categoryorder = "array",
              categoryarray = sp_summary$sp)

xform <- list(title="Independent detections")

# Capture rate
fig1 <- plot_ly(x = sp_summary$count, y = sp_summary$common_name, type = 'bar', orientation = 'h', showlegend=F) %>% 
 layout(yaxis = yform, xaxis=xform)

yform <- list(categoryorder = "array",
              categoryarray = sp_summary$sp,
              showticklabels=F)
xform <- list(title="Occupancy")


# Occupancy
fig2 <- plot_ly(x = sp_summary$occupancy, y = sp_summary$common_name, type = 'bar', orientation = 'h', showlegend=F) %>% 
 layout(yaxis = yform, xaxis=xform)

subplot(nrows=1,fig1, fig2, titleX = T) # We could stack them on top of one another using nrows=2

```

For the top twenty detected species, the treatment-specific capture rates are as follows:

```{r}
# Get species columns (everything except placename, days, plot_id, simple_treatment)
species_cols <- colnames(total_count)[!colnames(total_count) %in% c("placename", "days", "plot_id", "simple_treatment")]

# Pivot to long format
cam_long <- total_count %>%
  pivot_longer(cols = all_of(species_cols), 
               names_to = "species", 
               values_to = "count")

# Calculate capture rate per 100 days by species and treatment
capture_rates <- cam_long %>%
  group_by(species) %>%
  summarise(
    total_captures = sum(count, na.rm = TRUE),
    total_days = sum(days, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(capture_rate = (total_captures / total_days) * 100)

# Get top 20 species overall
top_20_species <- capture_rates %>%
  group_by(species) %>%
  summarise(total = sum(total_captures)) %>%
  arrange(desc(total)) %>%
  slice_head(n = 20) %>%
  pull(species)

# Get species order based on forest capture rate
#forest_order <- capture_rates %>%
#  filter(species %in% top_20_species, simple_treatment == "forest") %>%
#  arrange(desc(capture_rate)) %>%
#  pull(species)

# Add any species not detected in forest to the end
#forest_order <- c(forest_order, setdiff(top_20_species, forest_order))

# Swap species names with common names 
lookup <- img_sub[, c("sp", "common_name")]
lookup <- lookup[duplicated(lookup) == FALSE, ]

# Create common_name order based on forest_order
#common_name_order <- lookup$common_name[match(forest_order, lookup$sp)]

# Filter to top 20 and set factor levels
plot_data <- capture_rates %>%
  filter(species %in% top_20_species) %>%
  left_join(lookup, by = c("species" = "sp")) %>%
  mutate(
    species = factor(species),
    common_name = factor(common_name),
    #simple_treatment = factor(simple_treatment, 
    #                           levels = c("forest", "active_restoration", #"passive_restoration")),
    capture_rate_log = capture_rate + 0.01,
    text_color = ifelse(capture_rate < 5, "white", "black")
  )

# Create the heatmap with log scale
ggplot(plot_data, aes(x=1, y = common_name, fill = capture_rate_log)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(capture_rate, 1), color = text_color), size = 3) +
  scale_fill_viridis_c(option = "plasma", 
                       trans = "log10",
                       breaks = c(0.1, 1, 10, 100),
                       labels = c("0.1", "1", "10", "100"),
                       name = "Captures\nper 100 days") +
  scale_color_identity() +
  labs(
    y = "",
    title = "Top 20 species capture rates"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 14),
    plot.title = element_text(size = 16)
  )

```

## Spatial plot of capture rates

The overall capture rate by camera trap location. 

```{r}
 total_obs <- left_join(total_obs, locs)
 sp_list <- sp_summary$sp     
      # Calculate overall capture rate (sum of all species)
 overall_cr <- total_obs %>%
   mutate(
     total_detections = rowSums(select(., all_of(sp_list)), na.rm = TRUE),
     CaptureRate = total_detections / (days / 100)
     )
      
      # Calculate min and max for legend
      max_cr <- max(overall_cr$CaptureRate, na.rm = TRUE)
      min_cr <- min(overall_cr$CaptureRate, na.rm = TRUE)
      mid_cr <- min_cr + ((max_cr - min_cr)/2)
      
      # Create custom HTML legend
      legend_html <- paste0(
        '<div style="padding: 10px; background: white; border: 2px solid grey; border-radius: 5px;">',
        '<strong>All Species</strong><br>',
        '<strong>Capture Rate</strong><br>(per 100 days)<br><br>',
        '<svg width="120" height="80">',
        '<circle cx="15" cy="15" r="11" fill="rgba(0, 102, 204, 0.6)" />',
        '<text x="35" y="20" font-size="12">High: ', round(max_cr, 2), '</text>',
        '<circle cx="15" cy="40" r="7" fill="rgba(0, 102, 204, 0.6)" />',
        '<text x="35" y="45" font-size="12">Med: ', round(mid_cr, 2), '</text>',
        '<circle cx="15" cy="65" r="3" fill="rgba(0, 102, 204, 0.6)" />',
        '<text x="35" y="70" font-size="12">Low: ', round(min_cr, 2), '</text>',
        '</svg>',
        '</div>'
      )
      
      leaflet() %>%
        addProviderTiles(providers$Esri.WorldImagery) %>%
        addCircleMarkers(
          data = overall_cr,
          lng = ~longitude,
          lat = ~latitude,
          radius = ~(CaptureRate / max(CaptureRate, na.rm = TRUE) * 10) + 1,
          color = "#FF00E1",
          stroke = FALSE,
          fillOpacity = 0.8,
          popup = ~paste(placename, "<br>Total Detections:", total_detections, 
                         "<br>Capture Rate:", round(CaptureRate, 2))
        ) %>%   
        addScaleBar(position = "topleft") %>%
        addControl(html = legend_html, position = "bottomright")
    
```

Now the distribution per species



```{r}

img_subMammal<-img_sub %>% filter(class=="Mammalia")
Mammalspp<-unique(img_subMammal$sp)

img_subBird<-img_sub %>% filter(class=="Aves")

#################################################################

overall_cr_mammals <- overall_cr %>%
  select(placename, days, longitude, latitude, total_detections, CaptureRate, all_of(Mammalspp))

# Calculate capture rate for each species
for(species in Mammalspp) {
  overall_cr_mammals[[paste0(species, "_CR")]] <- 
    (overall_cr_mammals[[species]] / overall_cr_mammals$days) * 100
}

#################################################################

map <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addScaleBar(position = "topleft")

# Add individual species layers
for(i in 1:length(Mammalspp)) {
  cr_col <- paste0(Mammalspp[i], "_CR")
  species_cr_mammals <- overall_cr_mammals %>% 
    filter(get(cr_col) > 0)
    
  # Skip species with all zeros
  map <- map %>%
    addCircleMarkers(
      data = species_cr_mammals,
      lng = ~longitude,
      lat = ~latitude,
      radius = ~(get(cr_col) / max(get(cr_col), na.rm = TRUE) * 10) + 1,
      color = "lightyellow",
      stroke = FALSE,
      fillOpacity = 0.7,
      popup = ~paste(placename, "<br>", Mammalspp[i], ": ", round(get(cr_col), 2), 
                     " detections/100 days<br>Total detections:", get(Mammalspp[i])),
      group = Mammalspp[i],
      label = ~paste(Mammalspp[i], "-", round(get(cr_col), 2))
    )
}

# Add layer control
map <- map %>%
  addLayersControl(
    overlayGroups = c(Mammalspp[Mammalspp %in% names(overall_cr_mammals)]),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup(species_cols)  # Start with all species hidden, only show "All Mammals"

map


    
```



